---
title: "Unit6 Case Study Assignment"
author: "Aravind Veluchamy"
date: "June 21, 2016"
output: 
 html_document:
  keep_md: true
---
##This Case Study offers a glimpse into how the GDP Ranking of a country based on GDP Dollar Value in Millions stacks up against the Per capita income which is indiacted by the Income Group. The purpose of the case study is to do a preliminary investigation on relationship between GDP Value(from Rank) and Income group(Income per capita).The data has been obtained from world bank.
###The following  code chunk explains the downloading the GDP and Country data that was used in the study
###Loading Required Libraries
```{r}
library(downloader)
library(scales)
library(ggplot2)
```
###Downloading the data sets
```{r}
download("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv ",destfile = "GDP.csv")
download("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv  ",destfile = "country.csv")
GDPdata <- read.csv("GDP.csv",stringsAsFactors = FALSE,header=FALSE)
Country <- read.csv("country.csv",stringsAsFactors = FALSE,header=FALSE)
head(GDPdata)
str(GDPdata)
head(Country)
str(Country)
```

### The following code chunk explains cleaning the GDPdata and Country datasets and assigning Column Headers.
```{r}
GDPdata <- GDPdata[-c(1,2,3,4,5),]
GDPdata <- GDPdata[,-c(3,6,7,8,9,10)]
names(GDPdata) <-c("CountryCode","Rank","Country","DollarValue")
NoRank <- which(GDPdata$Rank=="")
GDPdata <-GDPdata[(-1*NoRank),]
NoDollarvalue <- which(GDPdata$DollarValue=="")
GDPdata <-GDPdata[(-1*NoDollarvalue),]
Country <- Country[,-c(10:31)]
Country <-Country[-c(1),]
names(Country) <-c("CountryCode" , "CountryName" , "IncomeGroup" , "Region" , "LendingCategory" , "OtherGroups" , "CurrencyUnit" , "LatestPopulationCensus" , "LatestHouseholdSurvey" )
```
###1.Merging Country and GDP data sets and number of records after merging is 189.
```{r}
GDPandCountry_Merged<- merge(GDPdata,Country,by ="CountryCode")
str(GDPandCountry_Merged)
```
###2.Sorting Merged data set by Rank and display the 13th country in Ascending order(US must be last).
####The 13th country in descending order is St.Kitts and Nevis.The original question was to sort data in ascending order ,but it was also mentioned that US should be last.Therefore I sorted the data in Descending order.
```{r}
GDPandCountry_Merged_sorted_Asc <- GDPandCountry_Merged
GDPandCountry_Merged_sorted_Asc <- transform(GDPandCountry_Merged_sorted_Asc,Rank=as.character(Rank))
GDPandCountry_Merged_sorted_Asc <- transform(GDPandCountry_Merged_sorted_Asc,Rank=as.integer(Rank))
GDPandCountry_Merged_sorted_Asc <-GDPandCountry_Merged_sorted_Asc[order(GDPandCountry_Merged_sorted_Asc$Rank,decreasing = TRUE),]
GDPandCountry_Merged_sorted_Asc[c(13),]
```
###3.What are the average GDP rankings for the "High income: OECD" and "High income: nonOECD" groups?
####The merged data set was divided into Highincome.NonOECD and Highincome.OECD using the subset function.The mean function was on the Rank column to obtain the Avg rankings for these two groups.
```{r}
Highincome.NonOECD <-subset(GDPandCountry_Merged,GDPandCountry_Merged$IncomeGroup=="High income: nonOECD")
Highincome.NonOECD <-transform(Highincome.NonOECD,Rank=as.character(Rank))
 Highincome.NonOECD <-transform(Highincome.NonOECD,Rank=as.integer(Rank))
mean(Highincome.NonOECD$Rank)
Highincome.OECD <-subset(GDPandCountry_Merged,GDPandCountry_Merged$IncomeGroup=="High income: OECD")
Highincome.OECD <-transform(Highincome.OECD,Rank=as.character(Rank))
 Highincome.OECD <-transform(Highincome.OECD,Rank=as.integer(Rank))
mean(Highincome.OECD$Rank)
```
###5.Cut the GDP ranking into 5 separate quantile groups. Make a table versus Income.Group. How many countries are Lower middle income but among the 38 nations with highest GDP?
####Firstly the Rank Column is converted into an integer type and then Dollar value undergoes a similar transformation.
####Eventhough Dollar Value is not used in this chunk the transformation is carried for plot function used in next chunk.
####Next the quantiles are plotted as 5 quantile groups.This is exactly how the data needs to cut in the next statement.
####Use the cut function to cut the data into five quantile groups and add a seperate column to capture the quantile group data falls into
####Plot a table of Income vs Group(represented by quantile range column)
```{r}
GDPandCountry_Merged_Quantiles <- GDPandCountry_Merged
GDPandCountry_Merged_Quantiles$Rank <- as.character(GDPandCountry_Merged_Quantiles$Rank)
GDPandCountry_Merged_Quantiles$Rank <- as.integer(GDPandCountry_Merged_Quantiles$Rank)
GDPandCountry_Merged_Quantiles$DollarValue = (gsub(",","",GDPandCountry_Merged_Quantiles$DollarValue))
GDPandCountry_Merged_Quantiles$DollarValue <- as.character(GDPandCountry_Merged_Quantiles$DollarValue)
GDPandCountry_Merged_Quantiles$DollarValue <- as.integer(GDPandCountry_Merged_Quantiles$DollarValue)
quantile(GDPandCountry_Merged_Quantiles$Rank,probs=seq(0,1,0.2))
GDPandCountry_Merged_Quantiles$QuantilesRange <- with(GDPandCountry_Merged_Quantiles,cut(GDPandCountry_Merged_Quantiles$Rank,breaks=quantile(GDPandCountry_Merged_Quantiles$Rank,probs=seq(0,1,by=0.2),na.rm=TRUE),include.lowest=TRUE))
Quantile <- table(GDPandCountry_Merged_Quantiles$IncomeGroup,GDPandCountry_Merged_Quantiles$QuantilesRange)
Quantile
```
###4.Plot the GDP for all of the countries. Use ggplot2 to color your plot by Income Group.
###Use the qplot function and then use the scale_y_continous to display the scales in numbers instead of exponential values
```{r}
CountryGDPinMils <- qplot(GDPandCountry_Merged_Quantiles$Rank,GDPandCountry_Merged_Quantiles$DollarValue,data=GDPandCountry_Merged_Quantiles,color=GDPandCountry_Merged_Quantiles$IncomeGroup)
CountryGDPinMils + scale_y_continuous(name="Country GDP in millions",labels=comma,limits=c(2,20000000),breaks=c(500000,1000000,3000000,6000000,9000000,12000000,15000000,18000000,20000000))+scale_x_discrete(name="Income Group",labels=c("High income: nonOECD","High income: OECD", "Low income","Lower middle income","Upper middle income"))
```
###Conclusion
####The GDP of a country alone is not a measure of Country's advancement or Quality of Life.Even though the country might have a High GDP it does not mean the population is having a good quality of life as the wealth might only benifit a certain section of society.